---
title: "SENTIMENT OF OCT 2016 VP DEBATE SPEECH"
author: "WW44SS"
date: "Oct 6, 2016"
output: 
    html_document:
        css: markdown7.css
        toc: true
        toc_depth: 1
        keep_md: true
---


###SUMMARY
 
This continuation of [BIAS AND CONTEXT IN PRESIDENTIAL DEBATE TEXTS](https://rpubs.com/ww44ss/Debate_Text), which focused on a "Bag of Words" approach to analyzing the text of Presidential Debates. 

This analysis shows a "Heat Map" of frequent words. It is not really a new analysis, but just a better way of visualizing the data.  

###DATA SOURCES AND METHODS
The text of the debate are downloaded from the [UCSB Presidency Project](http://www.presidency.ucsb.edu/debates.php). Transcripts were pasted into Apple Pages and stored as unformatted .txt files.  






```{r, "load local functions", echo=FALSE}

idem <-function(x){x}

## LOAD LOCAL FUNCTIONS
    directory <- "/Users/winstonsaunders/Documents/oct_2016_vp_debate/"
    ## load text of debate files
    source(paste0(directory, "load_debate_text.R"))
    ## select candidate text
    source(paste0(directory, "candidate_text.R"))
    ## turn candidate text in a Text Corpus
    source(paste0(directory, "text_tc.R"))
    ## selects candidate text and then turns that into a Text Corpus
    source(paste0(directory, "candidate_text_tc.R"))
    ## enable printing of multiple plots
    source(paste0(directory, "multiplot.R"))
    ## vector normialize
    source(paste0(directory, "vector.normalize.R"))

```

```{r, "load r libraries", echo=FALSE, warning=FALSE, message=FALSE}

     ## check for installed packages
    if(! ("tm" %in% rownames(installed.packages())))      {install.packages("tm")}
    if(! ("RWeka" %in% rownames(installed.packages())))   {install.packages("RWeka")}
    if(! ("SnowballC" %in% rownames(installed.packages()))) {install.packages("SnowballC")}
    if(! ("ggplot2" %in% rownames(installed.packages()))) {install.packages("ggplot2")}
    if(! ("xtable" %in% rownames(installed.packages())))  {install.packages("xtable")}
    if(! ("reshape2" %in% rownames(installed.packages())))  {install.packages("reshape2")}
    if(! ("dplyr" %in% rownames(installed.packages())))  {install.packages("dplyr")}

    library(dplyr)
    library(animation)
    library(tm)
    library(RWeka)
    library(SnowballC)
    library(ggplot2)
    library(xtable)
    library(reshape2)
    library(stringr)
    if(! ("stringr" %in% rownames(installed.packages())))  {install.packages("stringr")}
    if(! ("syuzhet" %in% rownames(installed.packages())))  {install.packages("syuzhet")}
    if(! ("viridis" %in% rownames(installed.packages())))  {install.packages("viridis")}
    library(stringr)
    library(viridis)
    library(syuzhet)
```

```{r, "find debate files in directory", echo=FALSE, warning=FALSE, message=FALSE}

    ## 
    ## .txt file detection
    

    directory <- "/Users/winstonsaunders/Documents/oct_2016_vp_debate/"
    file.list <- list.files(directory)
    
    cleaned.file <- file.list[grepl("_cleaned", file.list)]
    
    debate_all <- read.csv(paste0(directory, cleaned.file), stringsAsFactors = FALSE) %>% as_data_frame
  


    debate_all_sentiment <- debate_all %>% 
        mutate(sentiment = get_sentiment(debate_all$text, method = "afinn")) %>% group_by(name)

    i = 300
    ggplot(debate_all_sentiment %>% filter(X <= i), aes(x = X, y = sentiment, fill = name)) +
        geom_bar(stat = 'identity', alpha = 0.5) + 
        geom_smooth(se=FALSE, n=500, span = .1, size = 0.4) + 
        facet_grid(.~name) + theme_bw() + 
        xlim(range(debate_all_sentiment$X)) + 
        ylim(range(debate_all_sentiment$sentiment)) 
    
    
    saveGIF
    for (i in unique(debate_all_sentiment$X)) {
        print( ggplot(debate_all_sentiment %>% filter(X <= i), aes(x = X, y = sentiment, fill = name)) + 
            geom_bar(stat = 'identity', alpha = 0.5) + 
            geom_smooth(se=FALSE, n=500, span = .1, size = 0.4) + 
            facet_grid(.~name) + theme_bw() + 
            xlim(range(debate_all_sentiment$X)) + 
            ylim(range(debate_all_sentiment$sentiment))) 
    }
    }, interval = 0.2, movie.name = "ggplot2-brownian-motion.gif", ani.width = 800, ani.height = 400)

```
